### 2024-02-01 paper.tex
briefly explain overlap analysis

\subsection{Analysis of scBS datasets for benchmarks}
To analyze scBS data from \citet{kremer_scnmt}, single-cell CpG methylation reports from all conditions were first stored with \texttt{scbs prepare} and then smoothed with \texttt{scbs smooth} using the default bandwidth of 1000~bp.
\new{
This data was then analysed multiple times with different combinations of analysis methods, namely four ways to divide the genome into intervals, two ways to quantify methylation in these intervals, and four approaches for dimensionality reductions.
The following four sets of genomic intervals were used:
VMRs, obtained with \texttt{scbs scan} using the current default options (bandwidth $=2000$, stepsize $=10$, variance threshold $= 0.02$, minimum cell requirement $=6$);
adjacent tiles of 100~kb width;
promoter regions as defined by the \textpm2~kb domain around the TSS of coding genes;
and candidate cis-regulatory regions annotated by the ENCODE consortium \citep{encode2020expanded}.
Methylation was quantified either by averaging binary methylation values, or by calculating the shrunken mean of the residuals as described earlier.
For dimensionality reduction we used for different approaches. Three of them involved imputing missing values and applying PCA. Imputation with iterative PCA is described in the section above. We also imputed missing values with mean methylation levels for the given gene (or zeros for shrunken residuals). The same process was also combined with the selection of only high-quality features as is described in \citet{luo2017single}:}
We selected tiles spanning $\ge20$ CpG sites and with sequencing coverage in at least 70\% of cells.
We then imputed missing values by imputing the mean for each tile, centered the values, and performed PCA.
\new{
The last approach we tried was using MOFA+ (with default parameters) instead of PCA, which does not require to impute missing values. 
In all four cases we reduced the dimensionality of the input data to 15 PCs or MOFA factors. In some cases MOFA+ returned smaller number of factors, since some of the requested 15 had zero variance.
For visualization, these 15 PCs or MOFA factors were subjected to UMAP with parameters min\_dist $=0.2$, init $=$ ``spca''.
To flexibly adapt to data sets of different sizes, we set n\_neighbors  $=\frac{\sqrt{n}}{1.5}$ (rounded to the nearest integer), where $n$ is the total cell number.
}
The same analysis was repeated for \new{three additional scBS data sets \citep{luo2017single, bian2018single, argelaguet2019gastru}}, and for smaller data sets generated by randomly sub-sampling cells separately from \new{these} data sets.

>>>>

\subsection{Analysis of scBS datasets for benchmarks}
To analyze scBS data from \citet{kremer_scnmt}, single-cell CpG methylation reports from all conditions were first stored with \texttt{scbs prepare} and then smoothed with \texttt{scbs smooth} using the default bandwidth of 1000~bp.
\new{
This data was then analysed multiple times with different combinations of analysis methods, namely four ways to divide the genome into intervals, two ways to quantify methylation in these intervals, and four approaches for dimensionality reductions.
The following four sets of genomic intervals were used:
VMRs, obtained with \texttt{scbs scan} using the current default options (bandwidth $=2000$, stepsize $=10$, variance threshold $= 0.02$, minimum cell requirement $=6$);
adjacent tiles of 100~kb width;
promoter regions as defined by the \textpm2~kb domain around the TSS of coding genes;
and candidate cis-regulatory regions annotated by the ENCODE consortium \citep{encode2020expanded}.
Methylation was quantified either by averaging binary methylation values, or by calculating the shrunken mean of the residuals as described earlier.
For dimensionality reduction we used four different approaches. Three of them involved imputing missing values and applying PCA. Imputation with iterative PCA is described in the section above. We also imputed missing values with mean methylation levels for the given gene (or zeros for shrunken residuals). The same process was also combined with the selection of only high-quality features as is described in \citet{luo2017single}:}
We selected tiles spanning $\ge20$ CpG sites and with sequencing coverage in at least 70\% of cells.
We then imputed missing values by imputing the mean for each tile, centered the values, and performed PCA.
\new{
Lastly, we used MOFA+ with default parameters instead of PCA, which does not require imputation of missing values. 
In all four cases we reduced the dimensionality of the input data to 15 PCs or MOFA factors. In some cases MOFA+ returned a smaller number of factors, since some of the requested 15 had zero variance.
For visualization, these 15 PCs or factors were subjected to UMAP with parameters min\_dist $=0.2$, init $=$ ``spca''.
To flexibly adapt to data sets of different sizes, we set n\_neighbors  $=\frac{\sqrt{n}}{1.5}$ (rounded to the nearest integer), where $n$ is the total cell number.
}
The same analysis was repeated for \new{three additional scBS data sets \citep{luo2017single, bian2018single, argelaguet2019gastru}}, and for smaller data sets generated by randomly sub-sampling cells separately from \new{these} data sets.

\new{VMRs that intersect protein-coding gene bodies, CpG islands or cCREs were quantified by subtracting VMRs with at least one bp of overlap using \texttt{bedtools} \texttt{subtract} \texttt{-A} \citep{quinlan2010bedtools}.
}
